Evolution of income (equality) in Canada

### 2022-04-25 outputs/paper.Rmd
intro+data

---
title: "Resolving mental health issue could possibly reduce criminal behaviors in Canada"
author: 
  - Alicia Yang
thanks: "https://github.com/Alicia-y/Telling-stories-with-data-final-paper"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

library(tidyverse)
```

@@

In recent years, the Canadian society have been increasingly experiencing the issue of mental health. Research have shown that in every five randomly selected Canadians or every two randomly selected Canadians over 40 years old, one of them probably has a history of suffering from mental illness (). The rapid growth of this problem is likely to be caused by a combination of social and environmental factors, including the family history of the patients, drug abuse issue, stressful environment, etc. The spread of COVID pandemic is also being accused of accelerating the issue of mental health in Canada, because statistics have shown that compared to 2020, there is a 4% increase in the number of Canadians over 18 years old who have demonstrated the symptoms of mental illness in 2021(). Mental health issue can be a very serious problem, even though lots of people don't pay enough attention to it. If does not only negatively affect the feelings, mood and perception about life of the patients, but could also threaten the physical well-being of them and people around them in extreme cases. According to the Centre for Addiction and Mental Health, people with mental illness might have suicidal tendencies(depression) and are more likely to become criminals and victims of violent incidents/crimes. I'm interested in the validity of this claim, thus, this paper is aimed to investigate whether and how the mental health issue affects the reported crime rate in Canada.

@@

Our data is of penguins (Figure \@ref(fig:bills)).

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

Talk way more about it.

@@

>>>>

---
title: "Evolution of income (equality) in Canada"
author: 
  - Alicia Yang
thanks: "Code and data are available at: https://github.com/Alicia-y/Telling-stories-with-data-final-paper"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "People have claimed that individuals who are experiencing mental illness are more likely to be involved in violent incidents and crime activities either as victims or as criminals. To investigate more on this issue, this paper aims to explore how the mental health issue interacts with the reported crimes in Canada. By utilizing the data from Statistics Canada, we found that increase in mental health cases does have a positive correlation with the crime rates in Canada. More specifically, teenagers suffering from mental illness are much more likely to be involved in criminal activities. In addition, implications behind this correlation is discussed and possible solutions is proposed."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

library(tidyverse)
library(dplyr)
library(tidyr)
```

@@

In recent decades the world as a whole is developing fast economically, and one evidence of which is the rise in income level. As more countries become more developed in terms of their economy and eventually overcome their issues of poverty, their governments have gradually shift their goal to dealing with the problem of economic inequality. 
It concerns how the wealth and income of the population is distributed. Although we are generally living higher quality lives and getting higher incomes, another problem is being put forward by people who are claiming that the gap between the rich and poor is getting larger due to technological advancement, globalization, countries' governmental structure and policies, which betters off the people who are already wealthy by making it easy for them to seek for more opportunities, thus, snowballing their assets faster. Income disparity can be a huge issue. It makes the income per capita of a country less reflective of its people's true living quality, because income per capita does not tell the distribution of the income. The richest people may pull up the statistics about the average/median income level, but it doesn't mean the country is performing well economically as a whole. In addition to that, income inequality also have other potential negative effects. It could reduce the stability of the economy and increase the risk of financial crisis. It could also result in corruption, misallocation of resources as the rich is empowered economically, socially and even politically.
Canada is one of the country in the world that provides the most benefits and welfare, yet people still claimed that the economic inequality is rising along with the income over the past 2 decades(2). Therefore, I'm interested in whether this is a serious issue in Canada. If so, in which regions and what period of time is it a serious issue? And is it possible to predict how the disparity of income change in future?

The data used in this report is obtained from Open Government Data of Statistics Canada. The dataset is called "Upper income limit, income share and average income by economic family type and income decile", and it provides different kinds information about the income level of the households of Canada between 1976 and 2020 by the Canadian Income Survey. In this dataset, I'm interested in the year the data is from, the geographical location, the income decile so that we could observe the difference between the lowest and highest decile, and the actual value of the income. In section \@ref(data), a cleaned dataset is obtained

Associated with

Difficult and challenging to deal with

@@

The dataset, which is a summary of the income-related statistics in Canada between 1976 and 2020, is obtained from the Open Government Portal of Statistics Canada. The source of the data is from the Canadian Income Survey conducted annually. 
One possible bias came with the survey methodology was the over-coverage of the units that are not the target population and the under-coverage of

R [@citeR], and R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "knitr" [@citeknitr], "dplyr" [@citedplyr], and "kableExtra" [@citekableExtra] are utilized to create an extract of the cleaned dataset (Table \@ref(tab:dataextract)).

The variables of interest are

```{r}

raw_data %>%
  filter(
         `Income decile` == "Total deciles",
         `Geographical Location` == "Canada") %>%
  ggplot(aes(x=year, y=Income, colour = `Geographical Location`))+geom_point()+theme_minimal()
```

@@

https://www.nytimes.com/2020/07/01/opinion/economic-inequality-moral-philosophy.html#:~:text=Enough%20economic%20inequality%20can%20transform,society%20ruled%20by%20the%20rich.&text=Large%20inequalities%20of%20inherited%20wealth,and%20undercuts%20equality%20of%20opportunity.

(2)https://www.conferenceboard.ca/hcp/hot-topics/canInequality.aspx#:~:text=Income%20inequality%20in%20Canada%20has,income%20Canadians%20also%20lost%20share.
